### 2025-08-06 getting-started.qmd
20250806 - long and wide form

## Transform Data from Long to Wide {#sec-longToWide}

@@

Below, we widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable):

@@

## Transform Data from Wide to Long {#sec-wideToLong}

>>>>

## Long and Wide Form {#sec-longWideForm}

@@

### Transform Data from Long to Wide {#sec-longToWide}

Below, we transform the data from long form to wide form.
The long data are in "`player`-`season`-`week` form".
We widen the data by two variables (`season` and `week`), using `tidyverse`, so that the data are now in "`player` form" (where each row is uniquely identified by the `ID` variable):

@@

### Transform Data from Wide to Long {#sec-wideToLong}

### 2025-08-06 mixed-models.qmd
20250806 - long and wide form

When estimating mixed models, it is important to know the structure of the data.
Examples of data from nonindependent units include nested data, multiple membership data, and cross-classified data, as depicted in @fig-nestedDataStructures.

>>>>

For fitting mixed models, the data should be in [long form](#sec-longWideForm)—each player (or unit of interest) should have multiple rows based on the grouping structure.
When estimating mixed models, it is important to know the grouping structure of the data because the data structure influences the model syntax.
Examples of data from nonindependent units include nested data, multiple membership data, and cross-classified data, as depicted in @fig-nestedDataStructures.
